\section{Introduction}\label{sec:introduction}

Blockchain is a distributed network of computers (\emph{peers}).
Each peer receives transaction requests, that eventually packs
into new \emph{blocks}. These form a growing tree,
with each block pointing back to the previous one. Mining is the process of
block creation and includes a notion of chain quality to
decide which of the many tree branches eventually emerges
as the stable chain, on which all peers agree,
therefore reaching \emph{consensus}. Mining
is normally remunerated: block creators may
mint and earn a predetermined amount of cryptocurrency.

This very abstract description is explicitly generic \wrt{}
the notion of transaction, irrelevant in this paper:
a transaction is the execution of an operation
that modifies the state of an abstract machine.
For instance, it is a money transfer in Bitcoin~\cite{Nakamoto08} or
the execution of some code over shared objects in Ethereum~\cite{AntonopoulosW25}.
This description is also generic \wrt{} the mining algorithm. Namely,
Bitcoin pioneered proof of work~\cite{DworkN92},
while more modern blockchains mostly use
proof of stake~\cite{Kwon14}, and this paper focuses on proof
of space~\cite{Spoto25}.
In Bitcoin's proof of work,
minted blocks carry evidence of the performance of some computational
work for their creation. The more this work, the higher the
likelihood for the blocks to be integrated in the stable chain.
It is well-known that this induces high electricity costs for CPU
computation and data centers, with a high ecological impact (resource
consumption, e-waste). Moreover, this leads to the concentration of mining power
in countries where electricity cheaper, against the idea of full decentralization.
Proof of stake emerged as a reaction to this. Its idea is that
only a restricted set of peers (\emph{validators}) may mint blocks.
This makes mining cheaper, efficient and ecologically sustainable.
Non-validators may stake on the success of validators. That introduces some
decentralization, but
it is generally accepted that proof of stake is more centralized than proof of work.
Currently, it is the most used mining algorithm.
The idea of proof of space~\cite{AbusalahACKPR17,AtenieseBFG14,CohenP19,DziembowskiFKP15,ParkKFGAP18,RenD16,Spoto25} is that blocks carry evidence
of the allocation of a large data structure on disk for their creation.
The larger that structure, the higher the likelihood that
blocks get included in the stable chain. Proof of space is decentralized like
Bitcoin's proof of work and ecological like proof of stake.

This paper focuses on \emph{mobile mining}, that is,
on mining with a mainstream mobile device (typically, a phone or tablet).
In principle, this is appealing because:
%
\begin{itemize}
\item it leverages a huge existing hardware base, that otherwise remains unused
  for most of the time;
\item it introduces mining to a large portion of humanity, who owns a
  phone but not a powerful connected computer;
\item it can be easily updated, thanks to the automatic update system of mobile devices;
\item it simplifies the start-up of new blockchains, since it is easier to convince people
  to install an app than to run and maintain a software on an always connected computer.
\end{itemize}
%
Mobile mining has not been used yet because of obvious limitations
of mobile devices \wrt{} battery capacity and network bandwidth, in particular
in least favorite countries.

This paper makes the following contributions:
%
\begin{enumerate}
\item it presents mobile mining, its power and limitations;
\item it shows how proof of space emerges as a niche solution here, whose mining
  is less sensitive to these limitations;
\item it presents Mokaminter, our actual Android app for mobile mining with proof of space;
\item it reports experiments with Mokaminter, discussing its battery consumption,
  amount of exchanged data and earned cryptos on different devices.
\end{enumerate}
%
Mokaminter is a miner for mobile devices, it is not a blockchain
peer, it does not hold nor exchange blocks, it does not host the blockchain,
it does not use a blockchain database:
all that would be prohibitive on a mobile device.
A blockchain peer is an external computer that takes care of such aspects.
Mokaminter connects to a peer and mines for it.

The rest of this paper is organized as follows.
Sec.~\ref{sec:related_work} discusses related work.
Sec.~\ref{sec:mining} gives a uniform description of various mining algorithms and
discusses if such algorithms can be implemented on a mobile device.
Sec.~\ref{sec:pos} describes the proof of space mining algorithm implemented by Mokaminter.
Sec.~\ref{sec:mokaminter} describes the Mokaminter mobile app.
Sec.~\ref{sec:experiments} confirms its low battery cost,
its reduced bandwidth and the quality of its mining activity, in terms of earned crypto.
Sec.~\ref{sec:conclusion} concludes.
