\section{Mokamint's Proof of Space Mining Algorithm}\label{sec:pos}

Mokamint uses a mining algorithm based on a specific instance of proof of space,
derived from Signum's. The idea is relatively simple: each peer
allocates, on disk, its own \emph{plot}, \ie{}, a set of data structures called \emph{nonces}.
A nonce (Fig.~\ref{fig:waiting_time_nonce}) is an array of bytes, split into chunks called \emph{scoops}.
Whenever a peer needs to create a new block, to expand a chain of blocks $\xi$,
it computes a \emph{challenge} from $\xi$ that sends to its miner (Fig.~\ref{fig:mining}). The miner uses
that challenge to compute the minimal waiting time among that of the nonces
in its plot on disk (Fig.~\ref{fig:waiting_time_plot}).
The nonce that generates the minimal
waiting time is sent to the peer, that will selevct the minimal waiting time among those
of all its connected miners and use it to build the block that expands $\xi$.
That block can then be propagated to all peers, but only after the waiting time has elapsed,
or otherwise other peers might reject it as (still) invalid.

\begin{figure}[t]
  \begin{center}
    \begin{tikzpicture}[scale=1]

      \draw (-5.5,1) node {nonce};
      \draw [thick] (-11.5,0) rectangle +(1.75,0.5);
      \draw (-10.6,0.2) node {$\scoop_0$};
      \draw [thick] (-9.75,0) rectangle +(1.75,0.5);
      \draw (-8.85,0.2) node {$\scoop_1$};
      \draw [thick] (-8,0) rectangle +(1.75,0.5);
      \draw (-7.1,0.2) node {$\scoop_2$};
      \draw [dashed] (-6.25,0) rectangle +(1,0.5);
      \draw (-5.75,0.25) node {$\ldots$};
      \draw [thick,fill=verylightred] (-5.25,0) rectangle +(1.75,0.5);
      \draw (-4.375,0.2) node {$\scoop_i$};
      \draw [dashed] (-3.5,0) rectangle +(1,0.5);
      \draw (-3,0.25) node {$\ldots$};
      \draw [thick] (-2.5,0) rectangle +(1.75,0.5);
      \draw (-1.63,0.2) node {$\scoop_{4094}$};
      \draw [thick] (-0.75,0) rectangle +(1.75,0.5);
      \draw (0.13,0.2) node {$\scoop_{4095}$};

      \draw (-5.5,-0.5) node {the waiting time of the nonce \wrt{} a challenge $\langle i,\gamma\rangle$ is $\hash(\scoop_i\append\gamma)$};
%      \draw (-5.5, -0.9) node {normalized \wrt{} the current acceleration of the network};

    \end{tikzpicture}
  \end{center}
  \caption{A nonce is an array of bytes, split into chunks called scoops.
    The computation of the waiting time of a nonce \wrt{} the challenge
    $\langle i,\gamma\rangle$ selects the $i$th scoop of the nonce
    and computes a hash value that is interpreted as a numerical waiting time.}\label{fig:waiting_time_nonce}
\end{figure}

Let us formalize these mining algorithm now.
%
\begin{definition}[Challenge]\label{def:challenge}
  The set of \emph{challenges} is
  \[
  \Challenges=\left\{\langle\scoopnumber,\gamma\rangle\left|
  \begin{array}{l}
    0\le\scoopnumber<\numberofscoops\text{ and $\gamma$ is a hash for $h_\generation$}
  \end{array}
  \right.\right\}
  \]
  where $\numberofscoops$ is the number of scoops contained in a nonce
  (Tab.~\ref{tab:parameters} and Fig.~\ref{fig:waiting_time_nonce}).
  The $\gamma$ component of a challenge is said to be its \emph{generation}.
\end{definition}
%
The intuition is that $\gamma$ is a sort of digest of the history of the chain $\xi$ that
one wants to expand with a next block. Sec.~\ref{sec:blockchain_construction} will
show how a challenge is derived from $\xi$.

A \emph{scoop} is a pair of hashes. A \emph{nonce} is a \emph{progressive number} $p$, and
a list of $\numberofscoops>0$ scoops (see Fig.~\ref{fig:waiting_time_nonce}, where the progressive
number is not shown, for simplicity).
%
\begin{definition}[Scoop, Nonce]\label{def:scoop_nonce}
  The sets of \emph{scoops} and \emph{nonces} are
  \begin{align*}
    \Scoops&=\left\{\langle h_1,h_2\rangle\left|\begin{array}{l}
    h_1,h_2\text{ are hashes}
    \end{array}\right.\right\},\\
    \Nonces&=\left\{\langle p,\scoops\rangle\left|\begin{array}{l}
    p\in\nat\text{ and }\scoops\in\Scoops^\numberofscoops
    \end{array}\right.\right\}.
  \end{align*}
\end{definition}

\begin{figure}[t]
  \begin{center}
    \begin{tikzpicture}[scale=1]

      \draw (0.4,5.4) node {plot};

      \draw [thick] (-0.5,4.5) rectangle +(1.75,0.5);
      \draw (0.4,4.75) node {$\nonce_0$};
      \draw[->] (1.5,4.75) -- +(0.25,0);
      \draw (4.3,4.75) node {waiting time of $\nonce_0$ \wrt{} $\langle i,\gamma\rangle$};

      \draw [thick] (-0.5,4) rectangle +(1.75,0.5);
      \draw (0.4,4.25) node {$\nonce_1$};
      \draw[->] (1.5,4.25) -- +(0.25,0);
      \draw (4.3,4.25) node {waiting time of $\nonce_1$ \wrt{} $\langle i,\gamma\rangle$};

      \draw [thick,fill=verylightred] (-0.5,3.5) rectangle +(1.75,0.5);
      \draw (0.4,3.75) node {$\nonce_2$};
      \draw[->] (1.5,3.75) -- +(0.25,0);
      \draw (4.3,3.75) node {waiting time of $\nonce_2$ \wrt{} $\langle i,\gamma\rangle$};

      \draw [thick] (-0.5,3) rectangle +(1.75,0.5);
      \draw (0.4,3.25) node {$\nonce_3$};
      \draw[->] (1.5,3.25) -- +(0.25,0);
      \draw (4.3,3.25) node {waiting time of $\nonce_3$ \wrt{} $\langle i,\gamma\rangle$};

      \draw (0.4,2.375) node {$\vdots$};
      \draw [dashed] (-0.5,1.5) rectangle +(1.75,1.5);
      \draw (4.3,2.375) node {$\vdots$};

      \draw [thick] (-0.5,1) rectangle +(1.75,0.5);
      \draw (0.4,1.25) node {$\nonce_{n-3}$};
      \draw[->] (1.5,1.25) -- +(0.25,0);
      \draw (4.475,1.25) node {waiting time of $\nonce_{n-3}$ \wrt{} $\langle i,\gamma\rangle$};

      \draw [thick] (-0.5,0.5) rectangle +(1.75,0.5);
      \draw (0.4,0.75) node {$\nonce_{n-2}$};
      \draw[->] (1.5,0.75) -- +(0.25,0);
      \draw (4.475,0.75) node {waiting time of $\nonce_{n-2}$ \wrt{} $\langle i,\gamma\rangle$};

      \draw [thick] (-0.5,0) rectangle +(1.75,0.5);
      \draw (0.4,0.25) node {$\nonce_{n-1}$};
      \draw[->] (1.5,0.25) -- +(0.25,0);
      \draw (4.475,0.25) node {waiting time of $\nonce_{n-1}$ \wrt{} $\langle i,\gamma\rangle$};

      \draw [decorate,decoration={brace,amplitude=4pt},thick,xshift=0.1cm,yshift=0pt]
      (7.2,5) -- (7.2,0) node [midway,right,xshift=.2cm] {$\min$};

    \end{tikzpicture}
  \end{center}
  \caption{A plot is a set of nonces. Its waiting time
    \wrt{} a challenge $\langle i,\gamma\rangle$ is the minimal waiting time among those of its nonces
    \wrt{} $\langle i,\gamma\rangle$ (Fig.~\ref{fig:waiting_time_nonce}).
    For instance, it might be that of $\nonce_2$ in the picture.}\label{fig:waiting_time_plot}
\end{figure}

A \emph{prolog} identifies the parts involved in mining: the peer, the miner (Fig.~\ref{fig:genericity})
and the chain identifier of the blockchain for which mining is performed.
%
\begin{definition}[Prolog]\label{def:prolog}
  The set of \emph{prologs} is
  \[
  \Prologs=\left\{\langle\mathit{key}_\public^{\mathit{peer}},
  \mathit{key}_\public^{\mathit{miner}},\chainid\rangle\left|\begin{array}{l}
  \text{where $\mathit{key}_\public^{\mathit{peer}}\in\PublicKeys$
    and $\mathit{key}_\public^{\mathit{miner}}\in\PublicKeys$}\\
  \text{are the public keys of the key pairs}\\
  \text{that identify a blockchain peer and a miner, respectively,}\\
  \text{and $\chainid$ is the chain identifier of a blockchain network}
  \end{array}\right.\right\}.
  \]
\end{definition}
%
Prologs are relevant since they are used to mark nonces: each nonce is \emph{for a given prolog}.
This has a few consequences:
%
\begin{itemize}
\item Mokamint knows which peer and miner to reward for mining a new block: their public keys are available in the prolog
  of the nonce in the block;
\item a peer has no interest in stealing the nonce of another peer: the reward will go to the latter, anyway;
\item a miner cannot use the same plot for mining for more peers: the only valid peer is the one whose public key is written in the prolog of the nonces of the plot;
\item a miner cannot use the same plot for mining for more chains: the identifier of the only valid chain is written is the prolog of the nonces of the plot.
\end{itemize}
%
The last point solves a long-lasting issue in Signum's original algorithm, that only used $\mathit{key}_\public^{\mathit{peer}}$ as prologs.

Def.~\ref{def:nonce_construction} formalizes the construction of a nonce, given its progressive number and a prolog.
It uses a constant $\kappa>0$ (Tab.~\ref{tab:parameters})
to limit its computational cost, by avoiding hashing very large chunks of data.
The construction algorithm derives a sequence of hashes (steps~\ref{step:nonce_construction:seed}
and~\ref{step:nonce_construction:first_hash}) and constructs a \emph{final} hash from
all of them (step~\ref{step:nonce_construction:final_hash}) that uses to modify
the original sequence of hashes (step~\ref{step:nonce_construction:modify}).
Then it shuffles the sequence (step~\ref{step:nonce_construction:swap}), which intuitively
guarantees that, in order to compute a given scoop of the nonce
(\ie{}, a pair of consecutive hashes), the algorithm must be thoroughly executed.
%
\begin{definition}[$\nonce(p,\pi)$]\label{def:nonce_construction}
  Given $p\in\mathbb{N}$ and $\pi\in\Prologs$, let
  \[
  \nonce(p,\pi)=\langle p,\langle h_0,h_1\rangle\append
  \cdots\append\langle h_{2\cdot\numberofscoops-2},h_{2\cdot\numberofscoops-1}\rangle\rangle\in\Nonces,
  \]
  where the hashes $h_0,\ldots,h_{2\cdot\numberofscoops-1}$ are constructed as follows:
  %
  \begin{enumerate}
  \item\label{step:nonce_construction:seed} let $\seed=\pi.\mathit{key}_{\mathit{peer}}\append
    \pi.\mathit{key}_{\mathit{miner}}\append\pi.\mathit{chainid}\append p$;
  \item\label{step:nonce_construction:first_hash}
    for each $i$ from $2\cdot\numberofscoops-1$ to $0$, let
    \[
    h_i=\hash\left(\text{first $\kappa$ bytes of}\left(\left(\append_{i<j<2\cdot\numberofscoops}h_j\right)\append\seed\right)\right);
    \]
  \item\label{step:nonce_construction:final_hash}
    let
    $
    \finalhash=\hash\left(\left(\append_{0\le j<2\cdot\numberofscoops}h_j\right)\append\seed\right);
    $
  \item\label{step:nonce_construction:modify} for each $i$ from $0$ to $2\cdot\numberofscoops-1$, reassign
    $h_i$ to $h_i\xor\finalhash$;
  \item\label{step:nonce_construction:swap}
    for each odd $i$ from $1$ to $2\cdot\numberofscoops-1$, swap
    $h_i$ with $h_{2\cdot\numberofscoops-i}$.
  \end{enumerate}
\end{definition}
%
Let us discuss Def.~\ref{def:nonce_construction} now.
Step~\ref{step:nonce_construction:seed} starts a chain of hash computations that includes the prolog $\pi$
in the seed. As a consequence, the resulting nonces will be \emph{for $\pi$}: for instance,
the nonces of another peer or of another miner or for another chain identifier
will have a different $\pi$ and consequently
they will be different nonces, assuming that hash collisions are unlikely.
Step~\ref{step:nonce_construction:first_hash} is a loop that computes all $h_i$ from the seed.
It uses a hashing function $\hash$ that is expected to be expensive to compute, such
as shabal256 (Tab.~\ref{tab:parameters}): this makes it hard to recompute nonces and plots on-the-fly (which would
be a proof of work algorithm) rather than once and for all before mining.
Step~\ref{step:nonce_construction:final_hash} computes a digest $\finalhash$ of all $h_i$'s and
step~\ref{step:nonce_construction:modify} uses that digest to make each $h_i$ dependent on all others, so that
one must compute them all in order to compute each of them.
This is strengthened by the traditional use of $\oplus$ (the bitwise xor function) to minimize the information loss.
Step~\ref{step:nonce_construction:swap} makes each scoop include a $h_i$ with $i\ge\numberofscoops$,
so that smaller scoops are not easier to compute than the others. This avoids a potential
time/memory tradeoff attack that was exposed in the past for the original Signum algorithm
where step~\ref{step:nonce_construction:swap} did not exist:
it was possible to store only $\finalhash$ instead of the full nonce
and reconstruct the smaller scoops on demand. The resulting (mining power)/(space used)
ratio was proportionally higher than by storing the full nonces,
``at the price of having to compute a modest number of extra hashes''~\cite{ParkKFGAP18}.

Mokamint miners hold one or more plots on disk. Their initialization is performed
only once and offline, hence it is not part of the mining protocol.
Namely, a plot is a set of nonces constructed according to Def.~\ref{def:nonce_construction},
for a finite non-empty set of progressive numbers $P$ and
for a given prolog $\pi$, recorded in the plot.
%
\begin{definition}[Plot]\label{def:plot}
  The set of \emph{plots} is defined as
  \[
  \Plots=\left\{\left<\pi,\nonces\right>\left|\begin{array}{l}
  \pi\in\Prologs,\ \varnothing\not=P\subset\mathbb{N}\text{ is finite and }
  \nonces=\left\{\nonce(p,\pi)\left|\;p\in P\right.\right\}\text{ (Def.~\ref{def:nonce_construction}})
  \end{array}\right.\right\}.
  \]
\end{definition}

Given a challenge and a nonce, the latter has a \emph{value} that specifies how well
the nonce answers the challenge. The next definition formalizes the computation
in Fig.~\ref{fig:waiting_time_nonce}, up to the \emph{acceleration}, that will be considered later.
%
\begin{definition}[$\mathit{value}(\nonce,\challenge)$]\label{def:nonce_value}
  Let $\nonce\in\Nonces$ and $\challenge\in\Challenges$.
  The \emph{value of $\nonce$ \wrt{} $\challenge$} is
  \[
    \mathit{value}(\nonce,\challenge)=\hash(\nonce.\scoops[\challenge.i]\append\challenge.\gamma).
  \]
\end{definition}
%
Miners can use the value of their nonces to choose the best nonce for each given challenge,
\ie{}, the one with minimal value.
Note that only the $(\challenge.i)\mathit{th}$ scoop is used to compute that value.
That is, Mokamint consults only a small portion of the whole plot
file at each mining challenge, to spot the best nonce, that of minimal value.
That nonce could then be sent to the peer as answer for
the challenge (Fig.~\ref{fig:genericity}). But there are still two problems to solve:
%
\begin{enumerate}
\item the value in Def.~\ref{def:nonce_value} cannot be used as an absolute waiting time
  for the nonce yet, since it does not take into account the global mining power of the network:
  if the global mining power of the network increases (or decreases), then
  the minimal value of the nonces
  generated by the network tends to decrease (respectively, increase)
  and the block creation rate would not be
  fixed to $\beat$ (Tab.~\ref{tab:parameters}), on average.
  The value from Def.~\ref{def:nonce_value} must be normalized \wrt{} an
  \emph{acceleration} (Fig.~\ref{fig:waiting_time_nonce}),
  which is the inverse of Bitcoin's difficulty and has exactly its same function;
\item nonces are relatively big (around $262$ kbytes under the assumptions in the
  rightmost column of Tab.~\ref{tab:parameters}). Since nonces are stored in blockchain,
  one nonce inside each block, and since nonces are exchanged between miner and peer
  (Fig.~\ref{fig:genericity}), a more compact data structure is needed.
\end{enumerate}

\emph{Deadlines} are exactly this compact data structure that represents the nonce computed for a challenge.
Namely, a deadline carries the information needed
to reconstruct the nonce (progressive $p$ and prolog $\pi$, see Def.~\ref{def:nonce_construction}),
the value of that nonce (to avoid its repeated expensive recomputation through the formula
in Def.~\ref{def:nonce_value}), and
the challenge that that nonce was meant to answer. Therefore, in Fig.~\ref{fig:genericity}, the miner
will actually send back deadlines rather than nonces to the peer, which reduces the
data throughput. This is paramount if peer and miner are connected through a network rather
than being in the same physical machine.
%
\begin{definition}[Deadline]\label{def:deadline}
  The set of \emph{deadlines} is
  \[
  \Deadlines=\left\{
  \langle p,\pi,\mathit{value},\challenge\rangle
  \left|\begin{array}{l}
  \pi\in\Prologs,\ p\in\mathbb{N},\ \mathit{value}\text{ is a hash}\\
  \text{and }\challenge\in\Challenges
  \end{array}
  \right.
  \right\}.
  \]
  Deadlines are totally ordered by increasing value\footnote{Dealine's values are hashes, that can be interpreted as big-endian
  byte representations of natural numbers, that can be compared. Similarly, in Def.~\ref{def:deadline_waiting_time},
  the value of a deadline is interpreted as a natural number divident, in the same way. We do not formalize this explicitly in the
  notation, to keep it simpler.}.
\end{definition}
%
When a miner has selected its best nonce for a challenge, from a plot with prolog $\pi$,
and wants to send it to the peer,
it actually sends the much smaller deadline $\delta(\nonce,\pi,\challenge)$.
%
\begin{definition}[$\delta(\nonce,\pi,\challenge)$]\label{def:deadline_from_nonce}
  Given $\nonce\in\Nonces$, $\pi\in\Prologs$ and $\challenge\in\Challenges$, the
  \emph{deadline computed from $\nonce$ for $\pi$ and $\challenge$} is
  $\delta(\nonce,\pi,\challenge)=\langle\nonce.p,\pi,\mathit{value}(\nonce,\challenge),\challenge\rangle$.
\end{definition}
%
Conversely, when a peer receives a deadline from a miner, it must check that that deadline is \emph{valid}, \ie{},
that it is not lying about the value of the represented
nonce. It can do this by reconstructing that nonce and computing its value.
%
\begin{definition}[Deadline's validity]\label{def:deadline_validity}
  Let $\delta\in\Deadlines$. The deadline $\delta$ is \emph{valid} if and only if
  \[
  \delta.\mathit{value}=\mathit{value}(\nonce(\delta.p,\delta.\pi),\delta.\challenge).
  \]
\end{definition}

The value in a deadline becomes an absolute waiting
time as soon as the current acceleration of the network is taken into account.
%
\begin{definition}[Deadline's waiting time]\label{def:deadline_waiting_time}
  Given $\delta\in\Deadlines$ and an \emph{acceleration}
  $\alpha\in\mathbb{N}$ such that $\alpha>0$, the
  \emph{waiting time} of $\delta$ \wrt{} $\alpha$ is
  %
  \[
  \waitingtime(\delta,\alpha)
  %  =\betonat\left(\text{first $8$ bytes of}\left(\nattobe\left(\frac{\betonat(\delta.\mathit{value})}{\alpha}\right)\right)\right).
  =\frac{\delta.\mathit{value}}{\alpha}.
  \]
\end{definition}

\begin{algorithm}[t]
  \Input{a $\plot$, already computed on disk}
  \Input{the peer $P$ connected to $M$}
  \BlankLine
  \thread{$\mathsf{mineDeadline}(\plot,P)$}{
    \forever{}{
      \nl\label{step:deadline_mining:wait}
      wait to receive a $\challenge\in\Challenges$ from $P$\;
      %
      \nl\label{step:deadline_mining:minimize}
      identify the $\nonce\in\plot.\nonces$ that minimizes its value
      \wrt{} $\challenge$ (Def.~\ref{def:nonce_value})\;
      %
      \nl\label{step:deadline_mining:compute_deadline}
      send $\delta(\nonce,\plot.\pi,\challenge)$ to $P$ (Def.~\ref{def:deadline_from_nonce})\;
    }
  }
  \caption{The deadline mining algorithm, performed by each miner $M$ inside a dedicated thread.}
  \label{alg:deadline_mining}
\end{algorithm}

Alg.~\ref{alg:deadline_mining} is executed, inside a dedicated thread,
by each Mokamint miner $M$, connected to a peer $P$. It is an infinite loop that
waits for a challenge (step~\ref{step:deadline_mining:wait}), finds the best nonce
for that challenge, among those in the precomputed plot stored on its disk
(step~\ref{step:deadline_mining:minimize}), and sends back to $P$ the deadline
corresponding to that nonce (step~\ref{step:deadline_mining:compute_deadline}).
It is not required, but it is sensible that
the plot used in Alg.~\ref{alg:deadline_mining} has prolog
$\key_\public^P\append\key_\public^M\append\mathit{networkId}$, where
$\key_\public^P$ is the public key of the peer $P$, $\key_\public^M$ is the public key
of $M$ itself and $\mathit{networkId}$ is the chain identifier of the blockchain network
(Tab.~\ref{tab:parameters}). Namely, if another key is used instead of $\key_\public^P$ or if
another chain identifier is used instead of $\mathit{networkId}$, then the peer $P$ will
likely reject the deadline (as it will shown later, in Alg.~\ref{alg:block_mining},
steps~\ref{step:block_mining:check_deadline_chainid}
and~\ref{step:block_mining:check_deadline_peer}).
Furthermore, if another key is used instead of $\key_\public^M$, then $M$ would be working on behalf
of another miner, which is not in $M$'s interest.

As already observed, step~\ref{step:deadline_mining:minimize} of Alg.~\ref{alg:deadline_mining}, for each
received challenge,
consults only a small portion of the plot file to compute the best value, that is, the minimal waiting time
of the corresponding deadline
(it consults a single scoop for each nonce, see Fig.~\ref{fig:waiting_time_nonce}). Therefore, that step requires a fraction of
second only, even for large plot files. It follows that the life of a miner consists almost only in waiting
at step~\ref{step:deadline_mining:wait}, which explains the minimal energy-consumption of
Alg.~\ref{alg:deadline_mining}.
Namely, Fig.~\ref{fig:plot_and_mining_costs} reports the time for the creation of a plot file and the time for
using that file for mining, collected with Mokamint. Both grow linearly with the number of nonces in the plot file.
We recall that the goal of Signum's proof of space is to make the former expensive and the latter cheap. For instance, Fig.~\ref{fig:plot_and_mining_costs} shows that,
even with a powerful CPU, a plot of $4000$ nonces cannot
be recomputed at every block challenge, if the block creation rate is $20$ seconds.
That is, the nonces must be precomputed and stored on disk.

\begin{figure}[t]
  \begin{tabular}{|c||c|c|c|c|c|c|c|}
    \hline
    nonces & size & time for creation & time per challenge \\\hline\hline
    1000 & 256MBytes & 7.85s & 0.00078s \\\hline
    2000 & 512MBytes & 14.73s & 0.00156s \\\hline
    4000 & 1GByte & 31.47s & 0.00273s \\\hline
    8000 & 2GBytes & 64.80s & 0.00546s \\\hline
    16000 & 4GBytes & 130.99s & 0.01131s \\\hline
    32000 & 8GBytes & 257.71s & 0.02223s \\\hline
    64000 & 16GBytes & 554.25s & 0.04407s \\\hline
  \end{tabular}
  \caption{Time for the creation of plot files and for their use for mining:
    \emph{nonces} is the number of nonces in the plot file; \emph{size} is the size
    of the plot file; \emph{time for creation} is the time (in seconds) for creating the
    plot file (Def.~\ref{def:plot} and~\ref{def:nonce_construction});
    \emph{time for challenge} is the time (in seconds) to answer a challenge
    to compute a deadline, with that plot file (steps~\ref{step:deadline_mining:minimize}
    and~\ref{step:deadline_mining:compute_deadline} of Alg.~\ref{alg:deadline_mining}).
    Experiments have been performed on am Intel Core i7-11800H CPU.
    They can be verified by executing the \texttt{PlotProfiler} class
    of the \texttt{io-mokamint-plotter} module of~\cite{MokamintSourceCode}.
  }
  \label{fig:plot_and_mining_costs}
\end{figure}
